Analysis of cloud enhancement events in a 30-year record of global solar irradiance at Thessaloniki, Greece

### 2024-09-08 article/article.Rmd
Commit 2024-09-08 21:12

```{r loadtrends, echo=F, include=T, warning=FALSE, message=FALSE}
dailytrends <- data.table(
  read.csv("~/MANUSCRIPTS/02_enhancement/figures/Daily_trends_byYear.csv"))
dailyCE <- dailytrends[DATA == "ST_E_daily"]
yearlytrends <- data.table(
  read.csv("~/MANUSCRIPTS/02_enhancement/figures/Daily_trends_byYear_Proper.csv"))
yearlyCE <- yearlytrends[DATA == "ST_E_yearly"]

@@

In this study, we investigate the characteristics of global horizontal irradiance
enhancement events induced by clouds over Thessaloniki for the period 1994 -- 2023
using data recorded every one minute.  We identified the cloud enhancement (CE)
events by creating an appropriate cloud-free irradiance reference using a radiative
transfer model and aerosol optical depth data from a collocated Cimel sun photometer
and a Brewer spectrophotometer. We found a trend in CE events of
$`r plus_sign(round(yearlyCE[var=="GLB_diff.N",slope],0))`\pm
`r round(2*yearlyCE[var=="GLB_diff.N",slope.sd],0)`\,\text{cases}/\text{year}$,
and a trend in the corresponding irradiation of
$`r plus_sign(round(yearlyCE[var=="GLB_diff.sum",slope],1))`\pm
`r round(2*yearlyCE[var=="GLB_diff.sum",slope.sd],1)`\,\text{kJ}/\text{year}$.
The peak of the CE events was observed during May and June. CE events with duration
longer than 10\nobreakspace{}min are very rare
($<`r ceiling(100*ST_G0[,sum(GLB_ench.N>10)/.N])`\,\%$), with exceptions lasting
over an hour and up to 140 minutes.  Finally, we have detected enhancements above the
total solar irradiance at the top of the atmosphere for the same solar zenith angle
of up to
$209\,\text{W}/\text{m}^{2}$, with the $75\,\%$ of the cases below
$40\,\text{W}/\text{m}^{2}$. Most of these extreme events occur in spring -- early
summer, with a secondary peak in autumn.

@@

Following the application of the above discussed methodology to the entire dataset 
($`r round(ST_total$wattGLB.N / 10^6,1)`$ million of one-minute GHI measurements),
$`r round(100*ST_total$GLB_diff.N_pos/ST_total$wattGLB.N, 2)`\,\%$ were identified as CE events and
$`r round(100*ST_extreme_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ as ECE events.
The highest recorded GHI due to CE was
$`r round(MC[Typ == "Max Global", wattGLB],1)`\,\text{W}/\text{m}^2$ on
 `r strftime(MC[Typ == "Max Global", Date], "%d\\ %B %Y")` at a SZA of
$`r round(MC[Typ == "Max Global", SZA],1)`^\circ$ corresponding to OI of
$`r round(MC[Typ == "Max Global", GLB_diff],1)`\,\text{W}/\text{m}^2$ or
$`r round(100*MC[Typ == "Max Global", GLB_ench],1)`\,\%$ above the threshold.
The strongest CE event of
$`r round(100*MC[Typ == "Max Enha", GLB_ench],1)`\,\%$ above the 
cloud-free threshold was observed on 
`r strftime(MC[Typ == "Max Enh abs", Date], "%d\\ %B %Y")` at a SZA of
$`r round(MC[Typ == "Max Enh abs", SZA],1)`^\circ$ with a GHI value of 
$`r round(MC[Typ == "Max Enh abs", wattGLB],1)`\,\text{W}/\text{m}^2$ and a OI of
$`r round(MC[Typ == "Max Enh abs", GLB_diff],1)`\,\text{W}/\text{m}^2$.
Both cases are ECE events with
$`r round(MC[Typ == "Max Global", TSI_OI],1)`\,\text{W}/\text{m}^2$ and
$`r round(MC[Typ == "Max Enh abs", TSI_OI],1)`\,\text{W}/\text{m}^2$ above the
TSI at TOA for the same SZA, respectively.
In the following sections we are discussing the long-term trends and variability of
the CE events as well as of the corresponding OI and excess irradiation.

@@

Figure\nobreakspace{}\ref{fig:P-energy} shows the time series of the yearly number of
CE cases (each with duration of one minute), the yearly mean OI and the yearly excess
irradiation for the period 1994 -- 2023, together with corresponding linear trends.
To account for missing data, all three quantities have been divided with the fraction
of the valid GHI observations in each month.  Statistically significant (at the
$95\,\%$ confidence level) increasing trends appear for the yearly number of CE
events
($`r plus_sign(round(yearlyCE[var == "GLB_diff.N", slope], 0))`\pm `r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 0)`\,\text{cases}/\text{year}$)
and the excess irradiation
($`r plus_sign(round(yearlyCE[var == "GLB_diff.sum", slope], 1))` \pm `r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$),
while the the trend in the yearly mean OI is negligible
($`r plus_sign(round(yearlyCE[var == "GLB_diff.mean", slope], 2))`\pm `r round(2 * yearlyCE[var == "GLB_diff.mean", slope.sd], 1)`\,\text{W}/\text{m}^2/\text{year}$)
and of no statistical significance.
The average OI for the entire period is
$`r plus_sign(round(mean(ST_E_yearly$GLB_diff.mean), 1))`\pm `r round(sd(ST_E_yearly$GLB_diff.mean),1)`\,\text{W}/\text{m}^2$.
The interannual variability of the data about the trend lines is quite large.
Furthermore, the spread tends to increase with time (at least for the quantities of
panels b and c), suggesting a significant variability in cloud patterns over the
area, possibly associated to changes in climate.

>>>>

```{r loadtrends, echo=F, include=T, warning=FALSE, message=FALSE}
dailytrends <- data.table(
  read.csv("~/MANUSCRIPTS/02_enhancement/figures/Daily_trends_byYear.csv"))
dailyCE <- dailytrends[DATA == "ST_E_daily"]
yearlytrends <- data.table(
  read.csv("~/MANUSCRIPTS/02_enhancement/figures/Daily_trends_byYear_Proper.csv"))
## regular trends
yearlyCE <- yearlytrends[DATA == "ST_E_yearly"]

## trends with "complete data" !!!!!
yearlyCE <- yearlytrends[DATA == "ST_E_yearly_Complete"]

@@

In this study, we investigate the characteristics of global horizontal irradiance
enhancement events induced by clouds over Thessaloniki for the period 1994 -- 2023
using data recorded every one minute.  We identified the cloud enhancement (CE)
events by creating an appropriate cloud-free irradiance reference using a radiative
transfer model and aerosol optical depth data from a collocated Cimel sun photometer
and a Brewer spectrophotometer. We found a trend in CE events of
$`r plus_sign(round(yearlyCE[var=="GLB_diff.N",slope],0))`\pm
`r round(2*yearlyCE[var=="GLB_diff.N",slope.sd],0)`\,\text{cases}/\text{year}$,
and a trend in the corresponding irradiation of
$`r plus_sign(round(yearlyCE[var=="GLB_diff.sum",slope],1))`\pm
`r round(2*yearlyCE[var=="GLB_diff.sum",slope.sd],1)`\,\text{kJ}/\text{year}$.
The peak of the CE events was observed during May and June. CE events with duration
longer than 10\nobreakspace{}min are very rare
($<`r ceiling(100*ST_G0[,sum(GLB_ench.N>10)/.N])`\,\%$), with exceptions lasting
over an hour and up to 140 minutes.  Finally, we have detected enhancements above the
total solar irradiance at the top of the atmosphere for the same solar zenith angle
of up to
$204\,\text{W}/\text{m}^{2}$, with the $75\,\%$ of the cases below
$40\,\text{W}/\text{m}^{2}$. Most of these extreme events occur in spring -- early
summer, with a secondary peak in autumn.

@@

Following the application of the above discussed methodology to the entire dataset 
($`r round(ST_total$wattGLB.N / 10^6,1)`$ million of one-minute GHI measurements),
$`r round(100*ST_total$GLB_diff.N_pos/ST_total$wattGLB.N, 2)`\,\%$ were identified as CE events and
$`r round(100*ST_extreme_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ as ECE events.
The highest recorded GHI due to CE was
$`r round(MC[Typ == "Max Global", wattGLB],1)`\,\text{W}/\text{m}^2$ on
 `r strftime(MC[Typ == "Max Global", Date], "%d\\ %B %Y")` at a SZA of
$`r round(MC[Typ == "Max Global", SZA],1)`^\circ$ corresponding to OI of
$`r round(MC[Typ == "Max Global", GLB_diff],1)`\,\text{W}/\text{m}^2$ or
$`r round(100*MC[Typ == "Max Global", GLB_ench],1)`\,\%$ above the threshold.
The strongest CE event of
$`r round(100*MC[Typ == "Max Enh abs", GLB_ench],1)`\,\%$ above the 
cloud-free threshold was observed on 
`r strftime(MC[Typ == "Max Enh abs", Date], "%d\\ %B %Y")` at a SZA of
$`r round(MC[Typ == "Max Enh abs", SZA],1)`^\circ$ with a GHI value of 
$`r round(MC[Typ == "Max Enh abs", wattGLB],1)`\,\text{W}/\text{m}^2$ and a OI of
$`r round(MC[Typ == "Max Enh abs", GLB_diff],1)`\,\text{W}/\text{m}^2$.
Both cases are also ECE events with
$`r round(MC[Typ == "Max Global", TSI_OI],1)`\,\text{W}/\text{m}^2$ and
$`r round(MC[Typ == "Max Enh abs", TSI_OI],1)`\,\text{W}/\text{m}^2$ above the
TSI at TOA for the same SZA, respectively.
In the following sections we are discussing the long-term trends and variability of
the CE events as well as of the corresponding OI and excess irradiation.

@@

Figure\nobreakspace{}\ref{fig:P-energy} shows the time series of the yearly number of
CE cases (each with duration of one minute), the yearly mean OI and the yearly excess
irradiation for the period 1994 -- 2023, together with corresponding linear trends.
To account for missing data, all three quantities have been divided with the fraction
of the valid GHI observations in each month.  Statistically significant (at the
$95\,\%$ confidence level) increasing trends appear for the yearly number of CE
events
($`r plus_sign(round(yearlyCE[var == "GLB_diff.N", slope], 0))`\pm `r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 0)`\,\text{cases}/\text{year}$)
and the excess irradiation
($`r plus_sign(round(yearlyCE[var == "GLB_diff.sum", slope], 1))` \pm `r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$),
while the the trend in the yearly mean OI is negligible
($`r plus_sign(round(yearlyCE[var == "GLB_diff.mean", slope], 2))`\pm `r round(2 * yearlyCE[var == "GLB_diff.mean", slope.sd], 1)`\,\text{W}/\text{m}^2/\text{year}$)
and of no statistical significance.
The average OI for the entire period is
$`r plus_sign(round(mean(ST_E_yearly$GLB_diff.mean), 1))`\pm `r round(sd(ST_E_yearly$GLB_diff.mean),1)`\,\text{W}/\text{m}^2$.
The yearly excess irradiation due to CE events ranges between about $8$ and
$24\,\text{MJ}/\text{m}^2$. On average, it is about one half of the highest daily
irradiation of \~31 MJ/m^2^ under cloud-free conditions recorded at the summer
solstice in Thessaloniki.  Therefore, from solar energy perspective the yearly excess
irradiation is only a tiny fraction of the available solar energy. Although the
interannual variability of the mean OI is rather weak, the variability about the
trend lines is quite large for the number of CE cases and the excess irradiation
(panels b and c). For these two quantities the spread tends to increase with time,
suggesting a significant variability in cloud patterns over the area, possibly
associated to changes in climate.

The interannual variability of the data about the trend lines is quite large.
Furthermore, the spread tends to increase with time (at least for the quantities of
panels b and c), suggesting a significant variability in cloud patterns over the
area, possibly associated to changes in climate.
